D2F_exploration_fourway_and_GWAS

### 2020-05-27 R/D2F_exploration_fourway_and_gwas.Rmd
Manuscript additions looking for switchgrass floewring when exposed to shortening days of a specific length

J045.A SOURCE_ID_2 NSL 29896, Summer
```{r}
phe_parents <- phe_all %>%
  filter(PLANT_ID %in% c("DAC", "VS16", "J440.A", "J036.A", "J037.A")) 
  filter(PLANT_ID %in% c("AP13", "WBC", "J230.A")) 
  filter(PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "J440.A", "J440.B", "J230.A", "J292.A", "J045.A", "J036.A", "J037.A", "J022.B", "J210.A")) 
phe_F1 <- phe_all %>%
  filter(PLANT_ID %in% c("AxD", "VxW")) 
phe_F2 <- phe_all %>%
  filter(!(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "AxD", "VxW"))) 
```
Have J230.A, J440.A (2 sites: CLMB & PKLE)
For 440.A: J036.A at CLMB, KBSM, STIL; J037.A at 5 sites.
For 230.A: J022.B at 8 sites. J210.A at 4 sites. J660.A at two sites. Ok, ditch 660.A.

@@

```{r}
wea_gr_fl$SITE <- factor(wea_gr_fl$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
phe_all$SITE <- factor(phe_all$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
phe_parents$SITE <- factor(phe_parents$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
phe_F2$SITE <- factor(phe_F2$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))

@@

wea_gr_fl %>%
  dplyr::select(DOY, DYLN, SITE, RAIN) %>%
  unique() %>%
  filter(SITE != "OVTN") %>%
  ggplot(aes(x = DOY, y = DYLN)) + 
  geom_point() +
  facet_wrap(~SITE) + 
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = SUBPOP, y = 9), width = 0, height = 1, alpha = 0.1) +
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & WHERE == "FWCR"), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 6.5), width = 0, height = 1, alpha = 0.1) + 
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "FL50" & YEAR %in% c(2018,2019) & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = SUBPOP, y = 4), width = 0, height = 1, alpha = 0.1) +
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "FL50" & (YEAR %in% 2019 | YEAR == 2018 & SITE == "TMPL") & WHERE %in% c("FWCR")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 1.5), width = 0, height = 1, alpha = 0.1) +
  ylim(c(0,16)) +
  theme(legend.position = "right") +
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) + xlab("Day of year") + ylab("Daylength (hrs)") +
  geom_vline(xintercept = as_date(365/2-8, origin = "2018-12-31"), linetype = 2) +
  geom_text(aes(y = 1.5, x = as_date(52, origin = "2018-12-31"), label = "FWCR FL50")) +
  geom_text(aes(y = 4, x = as_date(52, origin = "2018-12-31"), label = "GWAS FL50")) +
  geom_text(aes(y = 6.5, x = as_date(242, origin = "2018-12-31"), label = "FWCR GR50")) +
  geom_text(aes(y = 9, x = as_date(242, origin = "2018-12-31"), label = "GWAS GR50")) + 
  geom_point(data = filter(wea_gr_fl, RAIN > 0 & SITE != "OVTN"), aes(y = RAIN/10), color = "blue", pch = 20)

@@

>>>>

J045.A SOURCE_ID_2 NSL 29896, Summer
```{r}
phe_F0_up <- phe_all %>%
  filter(PLANT_ID %in% c("DAC", "VS16", "J440.A", "J036.A", "J037.A")) 
phe_F0_low <- phe_all %>%
  filter(PLANT_ID %in% c("AP13", "WBC", "J230.A")) 
phe_F0 <- phe_all %>%
  filter(PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "J440.A", "J440.B", "J230.A", "J292.A", "J045.A", "J036.A", "J037.A", "J022.B", "J210.A")) 
phe_F1 <- phe_all %>%
  filter(PLANT_ID %in% c("AxD", "VxW")) 
phe_F2 <- phe_all %>%
  filter(!(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "AxD", "VxW"))) 
```
Have J230.A, J440.A (2 sites: CLMB & PKLE)
For 440.A: J036.A at CLMB, KBSM, STIL; J037.A at 5 sites.
For 230.A: J022.B at 8 sites. J210.A at 4 sites. J660.A at two sites. Ok, ditch 660.A.

@@

```{r}
wea_gr_fl$SITE <- factor(wea_gr_fl$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
phe_all$SITE <- factor(phe_all$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
phe_F0$SITE <- factor(phe_F0$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
phe_F2$SITE <- factor(phe_F2$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))

@@

wea_gr_fl %>%
  dplyr::select(DOY, DYLN, SITE, RAIN) %>%
  unique() %>%
  filter(SITE != "OVTN") %>%
  ggplot(aes(x = DOY, y = DYLN)) + 
  geom_point() +
  facet_wrap(~SITE) + 
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = SUBPOP, y = 9), width = 0, height = 1, alpha = 0.1) +
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & WHERE == "FWCR"), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 6.5), width = 0, height = 1, alpha = 0.1) + 
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "FL50" & YEAR %in% c(2018,2019) & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = SUBPOP, y = 4), width = 0, height = 1, alpha = 0.1) +
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "FL50" & (YEAR %in% 2019 | YEAR == 2018 & SITE == "TMPL") & WHERE %in% c("FWCR")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = as_factor(YEAR), y = 1.5), width = 0, height = 1, alpha = 0.1) +
  ylim(c(0,16)) +
  theme(legend.position = "right") +
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) + xlab("Day of year") + ylab("Daylength (hrs)") +
  geom_vline(xintercept = as_date(173, origin = "2018-12-31"), linetype = 2) +
  geom_text(aes(y = 1.5, x = as_date(52, origin = "2018-12-31"), label = "FWCR FL50")) +
  geom_text(aes(y = 4, x = as_date(52, origin = "2018-12-31"), label = "GWAS FL50")) +
  geom_text(aes(y = 6.5, x = as_date(242, origin = "2018-12-31"), label = "FWCR GR50")) +
  geom_text(aes(y = 9, x = as_date(242, origin = "2018-12-31"), label = "GWAS GR50")) + 
  geom_point(data = filter(wea_gr_fl, RAIN > 0 & SITE != "OVTN"), aes(y = RAIN/10), color = "blue", pch = 20)

@@

# Tie to previous research
## DYLN at FL plots
While the rate and extent of vegetative development is closely related to GDD, reproductive development is more tied to day-of-the-year, or photoperiod. Switchgrass is a short-day plant; it flowers when exposed to shortening days of a specific length (Benedict, 1940).
[so what is daylength of flowering for each plant_id * site – and is it the same across sites for individual plant_ids? Or different? If the same, this is the driver; if different, something else is the driver.]
```{r}
fl50_dyln <- phe_all %>%
  filter(PHE == "FL50") %>%
  left_join(wea_gr_fl, by = c("SITE", "YEAR", "MEAS" = "Day_of_year"))
fl50_dyln$SITE <- factor(fl50_dyln$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))

fl1_dyln <- phe_all %>%
  filter(PHE == "FL1") %>%
  left_join(wea_gr_fl, by = c("SITE", "YEAR", "MEAS" = "Day_of_year"))
fl1_dyln$SITE <- factor(fl1_dyln$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))

fl50_dyln %>%
  group_by(SUBPOP) %>%
  summarise(min_dyln = min(DYLN),
            max_dyln = max(DYLN))
max_dyln <- fl50_dyln %>%
  group_by(SITE) %>%
  summarise(min_dyln = min(DYLN, na.rm = TRUE),
            max_dyln = max(DYLN, na.rm = TRUE))
max_dyln <- wea_gr_fl %>%
  group_by(SITE) %>%
  summarise(min_dyln = min(DYLN, na.rm = TRUE),
            max_dyln = max(DYLN, na.rm = TRUE))

fl1_dyln %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
                               MEAS <= 172 ~ "DAYS_LENGTHENING",
                               TRUE ~ NA_character_)) %>%
  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("8X", "Atlantic")) %>% # , "Gulf", "Midwest"
  group_by(PLANT_ID) %>%
  add_tally() %>%
  filter(n == 8 | n > 13) %>%
  ggplot(aes(x = PLANT_ID, y = DYLN)) + 
  geom_point(aes(color = SITE, shape = SUBPOP)) + 
  facet_wrap(~FL_B4_SOL) +
  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right") + 
  scale_color_viridis(discrete = TRUE, direction = -1) +
  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))

fl1_dyln %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
                               MEAS <= 172 ~ "DAYS_LENGTHENING",
                               TRUE ~ NA_character_)) %>%
  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("Gulf", "Midwest")) %>% # , 
  group_by(PLANT_ID) %>%
  add_tally() %>%
  filter(n == 8 | n > 13) %>%
  ggplot(aes(x = PLANT_ID, y = DYLN)) + 
  geom_point(aes(color = SITE, shape = SUBPOP)) + 
  facet_wrap(~FL_B4_SOL) +
  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right", panel.spacing.x = unit(x = 5, units = "points")) + 
  scale_color_viridis(discrete = TRUE, direction = -1) +
  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))

fl1_dyln %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
                               MEAS <= 172 ~ "DAYS_LENGTHENING",
                               TRUE ~ NA_character_)) %>%
  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("Gulf", "Midwest")) %>% # , 
  group_by(PLANT_ID) %>%
  add_tally() %>%
  filter(n == 9 | n > 13) %>%
  ggplot(aes(x = PLANT_ID, y = DYLN)) +
  geom_violin(color = "grey", scale = "area") +
  geom_jitter(aes(color = SITE, shape = SUBPOP, fill = FL_B4_SOL), stroke = 1.2, size = 1.5, height = 0, width = 0.1) + 
  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right", panel.spacing.x = unit(x = 5, units = "points")) + 
  scale_color_viridis(discrete = TRUE, direction = -1) +
  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
  #facet_wrap(~SUBPOP) +
  scale_shape_manual(values = c(21, 24)) +
  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1)
  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))

fl1_dyln %>%
  mutate(FL_B4_SOL = case_when(MEAS > 173 ~ "DAYS_SHORTENING",
                               MEAS <= 172 ~ "DAYS_LENGTHENING",
                               TRUE ~ NA_character_)) %>%
  filter(YEAR == 2019 & WHERE == "GWAS" & SUBPOP %in% c("8X", "Atlantic")) %>% # , 
  group_by(PLANT_ID) %>%
  add_tally() %>%
  filter(n == 9 | n > 13) %>%
  ggplot(aes(x = PLANT_ID, y = DYLN)) +
  geom_violin(color = "grey", scale = "area") +
  geom_jitter(aes(color = SITE, shape = SUBPOP, fill = FL_B4_SOL), stroke = 1.2, size = 1.5, height = 0, width = 0.1) + 
  theme(axis.text.x = element_text(hjust = 1, angle = 45), legend.position = "right", panel.spacing.x = unit(x = 5, units = "points")) + 
  scale_color_viridis(discrete = TRUE, direction = -1) +
  geom_hline(data = max_dyln, aes(yintercept = max_dyln, color = SITE), linetype = 2) + 
  #facet_wrap(~SUBPOP) +
  scale_shape_manual(values = c(21, 24)) +
  scale_fill_viridis(option = "B", discrete = TRUE, direction = -1)
  geom_label(data = max_dyln, aes(x = 4, y = max_dyln, label = SITE))

```
Maybe also plot the max daylength at each site? 
Atlantic may be interesting to differentiate by ecotype or latitude of origin.

"The photoperiod response displayed within a given strain naturally differs with its latitude of origin and can be a point of exploitation in forage and biomass production systems."
```{r}

wea_gr_fl %>%
  dplyr::select(DOY, DYLN, SITE, RAIN) %>%
  unique() %>%
  filter(SITE != "OVTN") %>%
  ggplot(aes(x = DOY, y = DYLN)) + 
  geom_point() +
  facet_wrap(~SITE) + 
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = LATITUDE, y = 9), width = 0, height = 1, alpha = 0.25) +
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "GR50" & YEAR == 2019 & WHERE == "FWCR"), aes(x = as_date(MEAS, origin = "2018-12-31"), y = 6.5), width = 0, height = 1, alpha = 0.1) + 
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "FL50" & YEAR %in% c(2018,2019) & SUBPOP %in% c("Midwest", "Gulf")), aes(x = as_date(MEAS, origin = "2018-12-31"), color = LATITUDE, y = 4), width = 0, height = 1, alpha = 0.25) +
  geom_jitter(data = filter(phe_F2, SITE != "OVTN" & PHE == "FL50" & (YEAR %in% 2019 | YEAR == 2018 & SITE == "TMPL") & WHERE %in% c("FWCR")), aes(x = as_date(MEAS, origin = "2018-12-31"), y = 1.5), width = 0, height = 1, alpha = 0.1) +
  ylim(c(0,16)) +
  theme(legend.position = "right") +
  scale_color_viridis(end = 0.8, direction = -1) + xlab("Day of year") + ylab("Daylength (hrs)") +
  geom_vline(xintercept = as_date(365/2-8, origin = "2018-12-31"), linetype = 2) +
  geom_text(aes(y = 1.5, x = as_date(52, origin = "2018-12-31"), label = "FWCR FL50")) +
  geom_text(aes(y = 4, x = as_date(52, origin = "2018-12-31"), label = "GWAS FL50")) +
  geom_text(aes(y = 6.5, x = as_date(242, origin = "2018-12-31"), label = "FWCR GR50")) +
  geom_text(aes(y = 9, x = as_date(242, origin = "2018-12-31"), label = "GWAS GR50")) + 
  geom_point(data = filter(wea_gr_fl, RAIN > 0 & SITE != "OVTN"), aes(y = RAIN/10), color = "blue", pch = 20)
